---
title: "noco_climate.Rmd"
author: "Erika"
date: "4/23/2020"
output: html_document
---

April, 2020, in addition to being the heart of the COVID-19 outbreak, has been unseasonably cold!

That motivated me to play around with looking at climate data for the region.

I went to the [cornell climod2 website](http://climod2.nrcc.cornell.edu/)  and downloaded a bunch of data from the **Canton SE 4** station from a bunch of years (April 1, 1970- April 22, 2020).  

Prep the workspace
```{r}
rm(list = ls())
library(tidyverse)
```
And bring in the data

```{r}
DF<-read.csv("Data/canton4SE_1970_2020.csv", na.strings = c("M", " M"), stringsAsFactors = F) #brings in the M values (for "Missing") as NA
```
Now make the date column into a date, using lubridate.
```{r}
library(lubridate)
DF$Date<-ymd(DF$Date)
```
Now look at the data; need to be able to get from chr to numeric, but I suspect there are chrs dispersed throughout that need to be addressed.

T stands for "trace". Replace it throughout with 0.01 in the precipitation columns.

```{r}
fixem<-which(DF$Precipitation == " T")
DF$Precipitation[fixem]<- 0.01
fixem<-which(DF$Snowfall == " T")
DF$Snowfall[fixem]<-0.01
fixem<-which(DF$SnowDepth == " T")
DF$SnowDepth[fixem]<-0.01
```
Now make numeric
```{r}
DF$Precipitation<-as.numeric(DF$Precipitation)
DF$SnowDepth<-as.numeric(DF$SnowDepth)
DF$Snowfall<-as.numeric(DF$Snowfall)
#note - this makes strings containing the letter "A" or "S" into NA's which is good because S means subsequent and A means accumulated and thus those are summed rather than daily values. We'll ignore them.
```
Now get some summary data
```{r}
summary(DF)
```
Cool!  Now let's make some plots
```{r}
ggplot(DF, aes(Date))+
  geom_line(aes(Date, AvgTemperature),color = "black")
```
Add some date columns
```{r}
DF$Year<-year(DF$Date)
DF$Month<-month(DF$Date)
DF$Julian<-yday(DF$Date)
```
Assign each date to a season using meteorologoical seasons; According to meteorologists, 

Spring runs March through May (months 3 - 5)
Summer runs June through August (months 6 - 8)
Autumn runs September through November (month 9 - 11)
Winter runs December through February (monrha 12, 1 and 2)

```{r}
Spring<-which(DF$Month  == 3 | DF$Month == 4 | DF$Month == 5)
DF$Season<-NA
DF$Season[Spring]<-"Spring"
Summer<-which(DF$Month  == 6 | DF$Month == 7 | DF$Month == 8)
DF$Season[Summer]<-"Summer"
Fall<-which(DF$Month  == 9 | DF$Month == 10 | DF$Month == 11)
DF$Season[Fall]<-"Fall"
Winter<-which(DF$Month  == 12 | DF$Month == 1 | DF$Month == 2)
DF$Season[Winter]<-"Winter"
DF$Season<-as.factor(DF$Season)
```
Now make a new plot
```{r}
ggplot(DF, aes(Date, AvgTemperature, group = Season, color = Season))+
  geom_point()+
  geom_line()

```



